---
title: "ANA 515 Assignment 2, Loading, Saving, and Describing Data"
author: "Tomiris"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "##", prompt = F, message = F, warning = F
)
```

# Section 1: Description of the data 

I'm working with the Amazon dataset. This dataset contains over 10,000 different Amazon products and their information, user reviews, and so on for the year of 2020. The data is downloaded from https://www.kaggle.com/datasets/nguyenngocphung/10000-amazon-products-dataset in csv format. 

# Section 2: Reading the data into R

```{r}
# Using the fread() function from 'data.table' package
# Using it because the function is very fast at reading large volume of data 
# the amazon data set has so many rows and its size is more than 20 MB
# reading first 14 columns since they contain information
amazon <- data.table::fread("Amazon_Products.csv")[,1:14] 
```

# Section 3: Clean the data

```{r}
amazon <- amazon[,c(1:9)]  # taking the first 9 columns for convenience
# Renaming variables
colnames(amazon) <- c("ID","product","manufacturer","price","num_available","num_reviews","num_ans_ques","avg_rating","cat_and_sub_cat")
# Changing data types of some variables
amazon$price <- as.numeric(substring(amazon$price, 3))
amazon$avg_rating <- as.numeric(substring(amazon$avg_rating, 1, 4))
amazon$num_ans_ques <- as.numeric(amazon$num_ans_ques)
amazon$num_reviews <- as.numeric(amazon$num_reviews)
```

# Section 4: Characteristics of the data

This dataframe has `r nrow(amazon)` rows and `r ncol(amazon)` columns. The names of the columns and a brief description of each are in the table below:

Column Names  | Brief Descriptions
------------- | -------------
ID            | Unique ID of the product
product       | Product name
manufacturer  | Manufacturer name
price         | Price of the product in £ 
num_available | Number of products available in the stock
num_reviews   | Number of reviews
num_ans_ques  | Number of answered questions
avg_rating    | Average rating of reviews
cat_and_sub_cat | Amazon category and sub category

# Section 5: Summary statistics

```{r}
# Selecting the three columns
amazon_sub <- amazon[,c("price","num_reviews","avg_rating")]
# Creating a function that calculates the required summary statistics
summ <- function(df){
  su <- data.frame(Minimum = apply(df, 2, min, na.rm = TRUE),
                   Maximum = apply(df, 2, max, na.rm = TRUE),
                   Mean = apply(df, 2, mean, na.rm = TRUE),
                   No_Missing = apply(df, 2, function(x) sum(is.na(x))))
  return(su)  
}
# Assigning the result of summary statistics of the dataset
amazon_summary <- summ(amazon_sub)
knitr::kable(amazon_summary)
```


